
\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements
                                % to the table of contents

\comm{
This thesis work is the result of five years and two months long effort and it has been quite a professional and personal journey.
This dissertation is proof that I succeeded, a long awaited achievement that I first want to thank and to acknowledge myself for. \\
This accomplishment would not have been possible without the support of many people, inside and outside academia. I feel immensely grateful for them and their presence in my life. To those people I dedicate this final chapter of the thesis; the lenght shows how lucky am I to be able to count on this collection of people.\\

A dutiful recognition has to be made to my supervisor Didar Dobur for the opportunity to join the PhD program at UGent and for the given possibility to work on such interesting research topic. \\
I would like to express gratitude to the other professors in the experimental particle physics research group; Dirk, Michael and Archisman you have contributed to create a pleasant experience in a great scientific environment. 

I would like to personally thank one by one the exceptional physicists and colleagues I have worked with during the last five years.\\
First and foremost, in the long-lived Exotica group at CERN, I would like to thank Steven, Niki, Slava, Matthew, Juliette, Karri, Riccardo and Lesya for the fruitful discussions and essential advises we exchanged during various meetings and emails. CMS is a collaboration, no analysis is the result of the small group of analyzers listed in the CADI-line, but always the outcome of productive and brilliant ideas shared during coffee breaks and meeting for the seek of the best possible science. This is one of most valuable lessons that I had the privilege to learn at CERN. Thanks to Lesya for the inestimable work in the context of the prompt HNL analysis.\\
Special gratitude to the numerous members of the LHC LLP community; in particular thanks to Philippe, Marco, Jan, Richard, Martino, Carlos, James, Steven, Olivier, Brian and Jos\'e for your passion and role in the HNL hunting; with your discussions and expertise you have all significantly contributed to my education and my interest in long-lived HNLs. Thank you Marco, Richard and Jan for the insightful discussions and precious advises you were always inclined to give during the be.h meetings.  \\

Coming to the CMS UGent group, thanks to Daniele, you are one of the most knowledgeable people I had the opportunity to collaborate with and one of the few who was always willing to share insights and advise us.
Tom, without you I would still be writing "hello world" today, thank you for your expertise and the gigantic help on MC generation, SFs, ntuplizer, hiking tips, retro cruiser drives, etc, etc.
Thank you Kirill for your crucial contribution and your dedication during the decisive times of the LL-HNL analysis. 
Dear Joscha, you have been a breath of fresh air in this group with your vast knowledge and great enthusiasm for physics; I can say you have considerably contributed to the end of this PhD. \\
One of the most significant thank to my office mates: Willem, Tom and Illia; we were simply the best colleagues. Thank you Willem for sharing this incredible long journey; I honestly think that I made till the end because we were two and we knew what the other one was bearing at the time. Thank you for all your help from day one (the first year was a bit of a running-in period, though); you are really one of the most generous people when it comes to support and advices. I always felt I could trust you and rely on you and I am very grateful for all the endless and enriching conversations we had over the most disparate topics. I wish you the best and most successful career, just don't leave ham and yogurt on your desk, it is the worst business card. Finally thanks Illia for the nice discussions and the weird political conversations.\\
Big thank to my HNL buddy Basile, you always bring positive attitude and you have been a great companion during the numerous trips abroad for LLP and EXO workshops. Good luck!!!! Dear Marek, thank you for the incredible amount of coffee/beer breaks during the lockdown and the needed encouragements in the last months. I could always count on you bringing your peculiar grumpy attitude and funny jokes.  
 Moreover, I would like to genuinely thank the \emph{kids}, Liam, Gianny, Tu Tong and Luka for bringing new energy to the group and making the coffee breaks an essential part of the office days. I really appreciated the trust you gave to me in the last years. 
Finally, huge thank to the Center Office (Liam and Joscha) for making me write my thesis and immense gratitude to Gianny, Daniele, Liam, Joscha, Tu Tong, Luka, Basile and Sam for reading and correcting the manuscript! \\
I would like to thank the rest of the INW people who made me feel welcomed in this remote campus: Amrutha, Christos, Celine, Ianthe, Giel, Sander, Sam, Alexis, Andres, Ivan and Linda. \\

In the department of Physics and Astronomy I had a chance to meet incredible people who strongly contributed to my well being and made my PhD years memorable.  In order of ``discovery'', I would like to thank Alexis, Andres, Raul and Kajetan who turned the first years in Gent into a joyful and lively adventure making me feel at home in Gent. Thank you to Sam, Sander and Ianthe for the dancing Istanbul days. \\
Ianthe, I also wish I got to know you better sooner! But this is just the start of our new lives (finally!) and we will fill them with many shared future adventures, meetings and chats! Your support in the last years has been crucial in so many ways that i don't know how to describe them; I am mostly grateful for you always believing in me and for you trusting me in difficult moments. Finally, I can definitely say that has been such a fun time and more will come in the future. \\
Deniz, I don't even know where to start... You are my dear friend and I feel so so grateful for meeting you. You have contributed so profoundly to shape my last years; just to put things in perspective, I now like green color (before I hated it), I hear cars everywhere, I love Sonian forest, I know how to cook sea bass, I love Trieste, I respect animals, I eat better, I know great music, I evolve, I have sentences to help me in my struggles, I acknowledge my path, I know what an incredible mum is and I am an aunty of two extraordinary kids who made my life very precious. You are an exceptional and rare human being! I look forward to the new adventures of life we will share in the future, wherever that may be. A special thank to Jasper, Bora and Alp who welcomed me in their family with open arms.\\

Besides work, I spent amazing time with friends and family for which I am immensely grateful.\\
I would like to thank Fenna for the unconditional support, you took care of me with smile, laughs and sincere connection (and a lot of nice lockdown drinks). Thanks Evelin and Fenna for the great and silly time together, it was as needed as air at times. Thank you Isabella for our wild flights of imagination and your infinite ways to show me how much you love me. A big hug and thank to tiny Ana for being such an incredible example of energy and passion. \\
Juan... since that dinner at Aroy Aroy almost 6 years ago you have been my dearest friend here, I will not be able to thank you enough. I can not express how much you have changed and shape my path and my vision. I am pretty sure you are the person who challenges me the most and at the same time I truly feel you believe I will make things happen. Thank you for being by my side during amazing and memorable experiences. It has been a fun ride and a great privilege. \\
Thanks Kim, you have been one of the big suprises of the last few years. You have brought craziness, fun and openminded point of view that I did not even know I was looking for. Starting the two masters was definitly something audacious but I am happy I could say we happyly shared it.\\
My grateful thanks goes to Gent4Humanity people: Tom, Wilbert, Alex, Michelle, Marie, Alice, Nuria, Kee-Ran, James and Catalina; you gave me faith and hope. \\
Huge thanks to ICT4D penguins: Pietro, Enrico, Marko, Matteo, Gulten and Stephania. You were unconsciously the best moptivation and support during the last year. You made me believe I could turn this past experience in something useful and fullfilling. Thank Pietro for cheering me up but above all for helping me to see and define myself in the most authentic way.\\

Finally I want to thank the group of people who are the \emph{corner stone} of my life. \\
Grazie Michela e Benny per la copertina, per Frankie, per i giochi ma soprattutto per tutte le cene con cui mi accogliete ogni volta che sono in Italia facendomi sentire come se non me ne fossi mai andata. \\
Grazie Elena per aver creduto in me e in questo dottorato pi\'u di quanto ne fossi capace io stessa. \\
Grazie Ceci! Se penso a quanta strada abbiamo fatto da analisi1... sei una delle persone che mi ispira di pi\'u nella mia vita e che mi ha fatto credere potessi raggiungere tutto anche io! Sono sicura che diventerai il soggetto delle frasi future ``io ho una mia amica che salva il mondo, ed \'e la persona pi\'u figherrima che conosca''. Ci si vede a Ispra (e poi EU, ONU etc etc) Ceci!\\
Giulia Bruno, grazie... mi sembra superfluo e riduttivo scrivere qualunque cosa. Grazie per tutti i ``Martini/martiii'' che solo tu scrivi e che mi riempiono sempre il cuore, grazie per le tue milliardi di idee di cui mi rendi sempre partecipe, grazie per i sogni ad occhi aperti che condividiamo e grazie per farmi pensare al nostro futuro lontano condiviso nei porgetti che realizzaremo. \\
Un grazie enorme a Vale, Shana, Fra e Silvia. Siete il mio punto fisso che sono sicura mi accoglier\'a dopo qualunque avventura o idea impulsiva mi sia venuta. Grazie per aver avuto pazienza con me in questo eterno PhD, basta \'e finito. \\
Grazie Silvietta per essere la mia migliore amica... quella che mi conosce meglio di me stessa, con cui basta uno sguardo per avere tutto limpido e che c'\'e e basta, sempre. Tu ci sei e ci sei stata durante questi cinque anni facendomi sentire al sicuro come se vivessi qualche piano sopra di me, non hai idea quanta forza e certezza mi abbia dato.\\
Dulcis in fundo, un grazie immenso alla mia eclettica famiglia.\\
Grazie Cucone per i tuoi abbracci forti forti e per essere sempre la tua miss.\\
Grazie Tata... sei la mia fan numero uno e saperlo mi d\'a una forza di cui non sapevo di esserne in possesso. Grazie per aiutarmi a mettere sempre le cose in prospettiva o in un'altra prospettiva; grazie per avermi mostrato cosa voglia dire combattere tutti i giorni per cosa si crede giusto e grazie per avermi insegnato ad ascoltarmi anche quando tutto sembrava portarmi via. Di sicuro non avrei finito questa tesi senza di te, magari ti posso aggiungere come co-author... \\
Infine grazie pap\'a e mamma per avermi sostenuto in questi cinque anni lontano e per essere stati sempre curiosi di questo dottorato.\\
Un abbraccio a Mari.





%%%%%%%%%%%%%%%%%%%%%
Five years and two months of effort have gone into this thesis, which has been a challenging and exciting experience from a professional and personal standpoint.
This dissertation is proof that I succeeded, a long awaited achievement that I first want to thank and to acknowledge myself for. \\
Many people, both within and outside the academy, played a major role in making this achievement possible. I am deeply grateful to them and their contributions. This final chapter is dedicated to those people; its length reflects the luck I've had in being able to work with them.

A dutiful recognition has to be made to my supervisor Didar Dobur for the opportunity to join the PhD program at UGent and for the given possibility to work on such an interesting research topic. \\
I would like to express gratitude to the other professors in the experimental particle physics research group; Dirk, Michael and Archisman you have contributed to create a pleasant experience in a great scientific environment. 

I would like to personally thank one by one the exceptional physicists and colleagues I have worked with during the last five years.\\
First and foremost, in the long-lived Exotica group at CERN, I would like to thank Steven, Niki, Slava, Matthew, Juliette, Karri, Riccardo and Lesya for the fruitful discussions and essential advice we exchanged during various meetings and emails. A CMS analysis is never the result of the small group of analyzers listed on the CADI-line but is always a result of creative and productive ideas shared during coffee breaks and meetings for the pursuit of the best science. It was one of the most valuable lessons I learned at CERN. Thanks to Lesya for the inestimable work in the context of the prompt HNL analysis.\\
Special gratitude to the numerous members of the LHC LLP community; in particular thanks to Philippe, Marco, Jan, Richard, Martino, Carlos, James, Steven, Olivier, Brian and Jos\'e for your passion and role in the HNL hunting; with your discussions and expertise you have all significantly contributed to my education and my interest in long-lived HNLs. Thank you Marco, Richard and Jan for the insightful discussions and precious advice you were always inclined to give during the be.h meetings.  \\

Coming to the CMS UGent group, Daniele is one of the most knowledgeable individuals I have ever worked with, and one of the few who has always offered advice and insight.
Tom, without you I would still be writing "hello world" today, thank you for your expertise and the gigantic help on MC generation, SFs, ntuplizer, hiking tips, retro cruiser drives, etc, etc.
Thank you Kirill for your crucial contribution and your dedication during the decisive times of the LL-HNL analysis. 
Dear Joscha, you have been a breath of fresh air in this group with your vast knowledge and great enthusiasm for physics; I am sure you have significantly contributed to the end of this PhD. \\
One of the most significant thanks to my office mates: Willem, Tom and Illia; we were simply the best colleagues. Thank you Willem for sharing this incredibly long journey with me. I honestly believe that I made it to the end because we were two and we knew what the other was bearing at the time. You are really one of the most generous people when it comes to providing support and advice. I really appreciate all your help since day one (although the first year was a bit of a stumbling block). I've always felt I could trust and rely on you and I'm thankful for all the many and enriching conversations we've had over the years disparate topics. I wish you the best and most successful career, just don't leave ham and yogurt on your desk, it is the worst business card. Finally thanks Illia for the nice discussions and the weird political conversations.\\
Big thanks to my HNL buddy Basile, you always bring a positive attitude and you have been a great companion during the numerous trips abroad for LLP and EXO workshops. Good luck!!!! Dear Marek, thank you for the incredible amount of coffee/beer breaks during the lockdown and the needed encouragement in the last months. Your peculiar grumpy attitude and funny jokes could always be counted on.  
Moreover, I would like to genuinely thank the \emph{kids}, Liam, Gianny, Tu Tong and Luka for bringing new energy to the group and making the coffee breaks an essential part of the office days. I really appreciated the trust you gave to me in the last years. 
Finally, a huge thanks to the Center Office (Liam and Joscha) for making me write my thesis and immense gratitude to Gianny, Daniele, Liam, Joscha, Tu Tong, Luka, Basile and Sam for reading and correcting the manuscript! \\
I would like to thank the rest of the INW people who made me feel welcomed in this remote campus: Amrutha, Christos, Celine, Ianthe, Giel, Sander, Sam, Alexis, Andres, Ivan and Linda. \\

It was in the department of Physics and Astronomy that I met incredible people who greatly contributed to my well-being and made my PhD years memorable.  In order of ``discovery'', I would like to thank Alexis, Andres, Raul and Kajetan who turned the first years in Gent into a joyful and lively adventure making me feel at home in Gent. Thank you to Sam, Sander and Ianthe for the dancing Istanbul days. \\
Ianthe, I also wish I got to know you better sooner! It's just the beginning of our new lives (finally!) and we look forward to many adventures, meetings, and chats to follow! Your support in the last years has been crucial in so many ways that I don't know how to describe them; I am mostly grateful for you always believing in me and for trusting me in difficult moments. Last but not least, I can confirm that it has been a great time, and more is on the way. \\
Deniz, I don't even know where to start... You're such a dear friend to me, and I'm so grateful to have met you. You have contributed so profoundly to shape my last years; just to put things in perspective, I now like green color (before I hated it), I hear cars everywhere, I love Sonian forest, I know how to cook sea bass, I love Trieste, I respect animals, I eat better, I know great music, I evolve, I have sentences to help me in my struggles, I acknowledge my path, I know what an incredible mum is and I am an aunty of two extraordinary kids who made my life very precious. You are an exceptional and rare human being! I look forward to the new adventures of life we will share in the future, wherever that may be. A special thank to Jasper, Bora and Alp who welcomed me in their family with open arms.\\

Besides my work, I spent amazing time with family and friends, for which I am eternally grateful.
Thanks to Fenna for the unconditional support, you took care of me with smiles, laughs and sincere connection (and lots of lockdown drinks). Thanks Evelin and Fenna for a great and silly time together. It was necessary at times. Isabella, thank you for all our wild flights of imagination and your endless ways to show me how much you love me. Hugs and gratitude to tiny Ana for being such an inspiring example of energy and passion. \\
Juan... ever since that dinner at Aroy Aroy almost 6 years ago, you have been my dearest friend here, I cannot thank you enough. You have shaped my path and my vision to an extent that I cannot put into words. It's pretty clear you are the person who challenges me the most, and at the same time you seem to believe I will accomplish what I set out to do. Thank you for staying by my side during remarkable and memorable experiences. I enjoyed our time together greatly. \\
Kim, you've been one of the biggest surprises of the past few years. You have brought craziness, fun, and an open-minded point of view that I didn't even know I was looking for. Starting the two masters was something audacious but I am happy I could say we happily shared it.\\
My thanks go to Gent4Humanity people: Tom, Wilbert, Alex, Michelle, Marie, Alice, Nuria, Kee-Ran, James, and Catalina; you gave me faith and hope. \\
Huge thanks to ICT4D penguins: Pietro, Enrico, Marko, Matteo, Gulten, and Stephania. You were unconsciously the best motivation and support during the last year. You made me believe I could turn this experience into something useful and fulfilling. Thank Pietro for cheering me up but above all for helping me to see and define myself most authentically.\\

Finally I want to thank the group of people who are the \emph{cornerstone} of my life. \\
Grazie Michela e Benny per la copertina, per Frankie, per i giochi ma soprattutto per tutte le cene con cui mi accogliete ogni volta che sono in Italia facendomi sentire come se non me ne fossi mai andata. \\
Grazie Elena per aver creduto in me e in questo dottorato pi\'u di quanto ne fossi capace io stessa. \\
Grazie Ceci! Se penso a quanta strada abbiamo fatto da analisi1... sei una delle persone che mi ispira di pi\'u nella mia vita e che mi ha fatto credere potessi raggiungere tutto anche io! Sono sicura che diventerai il soggetto delle frasi future ``io ho una mia amica che salva il mondo, ed \'e la persona pi\'u figherrima che conosca''. Ci si vede a Ispra (e poi EU, ONU etc etc) Ceci!\\
Giulia Bruno, grazie... mi sembra superfluo e riduttivo scrivere qualunque cosa. Grazie per tutti i ``Martini/martiii'' che solo tu scrivi e che mi riempiono sempre il cuore, grazie per le tue miliardi di idee di cui mi rendi sempre partecipe, grazie per i sogni ad occhi aperti che condividiamo e grazie per farmi pensare al nostro futuro lontano condiviso nei progetti che realizzeremo. \\
Un grazie enorme a Vale, Shana, Fra e Silvia. Siete il mio punto fisso che sono sicura mi accoglier\'a dopo qualunque avventura o idea impulsiva mi sia venuta. Grazie per aver avuto pazienza con me in questo eterno PhD, basta \'e finito. \\
Grazie Silvietta per essere la mia migliore amica... quella che mi conosce meglio di me stessa, con cui basta uno sguardo per avere tutto limpido e che c'\'e e basta, sempre. Tu ci sei e ci sei stata durante questi cinque anni facendomi sentire al sicuro come se vivessi qualche piano sopra di me, non hai idea quanta forza e certezza mi abbia dato.\\
Dulcis in fundo, un grazie immenso alla mia eclettica famiglia.\\
Grazie Cucone per i tuoi abbracci forti forti e per essere sempre la tua miss.\\
Grazie Tata... sei la mia fan numero uno e saperlo mi d\'a una forza di cui non sapevo di esserne in possesso. Grazie per aiutarmi a mettere sempre le cose in prospettiva o in un'altra prospettiva; grazie per avermi mostrato cosa voglia dire combattere tutti i giorni per cosa si crede giusto e grazie per avermi insegnato ad ascoltarmi anche quando tutto sembrava portarmi via. Di sicuro non avrei finito questa tesi senza di te, magari ti posso aggiungere come co-author... \\
Infine grazie pap\'a e mamma per avermi sostenuto in questi cinque anni lontano e per essere stati sempre curiosi di questo dottorato.\\
Un abbraccio a Mari.

%%%%%%%%%%%%%%%%%%%%%
}






















































\vspace{1cm}


{\footnotesize\textsf{``Si stava così bene tutti insieme, così bene, che qualcosa di straordinario doveva pur accadere.
Bastò che a un certo momento lei dicesse: -Ragazzi, avessi un po’ di spazio, come mi piacerebbe
farvi le tagliatelle!- E in quel momento tutti pensammo allo spazio che avrebbero occupato le
tonde braccia di lei muovendosi avanti e indietro con il mattarello sulla sfoglia di pasta, il petto
di lei calando sul gran mucchio di farina e uova che ingombrava il largo tagliere mentre le sue
braccia impastavano impastavano, bianche e unte d’olio fin sopra al gomito; pensammo allo
spazio che avrebbero occupato la farina, e il grano per fare la farina, e i campi per coltivare il
grano, e le montagne da cui scendeva l’acqua per irrigare i campi, e i pascoli per le mandrie di
vitelli che avrebbero dato la carne per il sugo; allo spazio che ci sarebbe voluto perché il Sole
arrivasse con i suoi raggi a maturare il grano; allo spazio perché dalle nubi di gas stellari il Sole si
condensasse e bruciasse; alle quantità di stelle e galassie e ammassi galattici in fuga nello spazio
che ci sarebbero volute per tener sospesa ogni galassia ogni nebula ogni sole ogni pianeta, e nello
stesso tempo del pensarlo questo spazio inarrestabilmente si formava, nello stesso tempo in cui la
signora Ph(i)Nko pronunciava quelle parole: -...le tagliatelle, ve’, ragazzi!- il punto che conteneva
lei e noi tutti s’espandeva in una raggiera di distanze d’anni-luce e secoli-luce e miliardi di millenniluce,
e noi sbattuti ai quattro angoli dell’universo (il signor Pbert Pberd fino a Pavia), e lei dissolta
in non so quale specie d’energia luce calore, lei signora Ph(i)Nko, quella che in mezzo al chiuso
nostro mondo meschino era stata capace d’uno slancio generoso, il primo, “Ragazzi, che tagliatelle
vi farei mangiare!”, un vero slancio d’amore generale, dando inizio nello stesso momento al
concetto di spazio, e allo spazio propriamente detto, e al tempo, e alla gravitazione universale, e
all’universo gravitante, rendendo possibili miliardi di miliardi di soli, e di pianeti, e di campi di
grano, e di signore Ph(i)Nko sparse per i continenti dei pianeti che impastano con le braccia unte
e generose infarinate, e lei da quel momento perduta, e noi a
rimpiangerla.''}

\enquote{\itshape We got along so well all together, so well that something extraordinary was bound 
to happen. It was enough for her to say, at a certain moment: "Oh, if I only had some 
room, how I'd like to make some noodles for you boys!" And in that moment we all 
thought of the space that her round arms would occupy, moving backward and forward 
with the rolling pin over the dough, her bosom leaning over the great mound of flour and 
eggs which cluttered the wide board while her arms kneaded and kneaded, white and 
shiny with oil up to the elbows; we thought of the space that the flour would occupy, and 
the wheat for the flour, and the fields to raise the wheat, and the mountains from which 
the water would flow to irrigate the fields, and the grazing lands for the herds of calves 
that would give their meat for the sauce; of the space it would take for the Sun to arrive 
with its rays, to ripen the wheat; of the space for the Sun to condense from the clouds of 
stellar gases and bum; of the quantities of stars and galaxies and galactic masses in flight 
through space which would be needed to hold suspended every galaxy, every nebula, 
every sun, every planet, and at the same time we thought of it, this space was inevitably 
being formed, at the same time that Mrs. Ph(i)Nk 0 was uttering those words: "... ah, 
what noodles, boys!" the point that contained her and all of us was expanding in a halo of 
distance in light-years and light-centuries and billions of light-millennia, and we were 
being hurled to the four corners of the universe (Mr. Pber* Pber d all the way to Pavia), and 
she, dissolved into I don't know what kind of energy-light-heat, she, Mrs. Ph(i)Nk 0 , she 
who in the midst of our closed, petty world had been capable of a generous impulse, 
"Boys, the noodles I would make for you!," a tme outburst of general love, initiating at 
the same moment the concept of space and, properly speaking, space itself, and time, and 
universal gravitation, and the gravitating universe, making possible billions and billions 
of suns, and of planets, and fields of wheat, and Mrs. Ph(i)Nk 0 s, scattered through the 
continents of the planets, kneading with floury, oil-shiny, generous amis, and she lost at 
that very moment, and we, mourning her loss.  }\bigbreak
}

\hfill Italo Calvino "Le Cosmicomiche", 1965.
\end{acknowledgements}